\speculake: Hiding Malware in Speculative Execution

### 2018-05-07 paper/appendix_spasm.tex
partial related work, and bib fixes across paper
Read through - minor updates

\section{SPASM Instruction Set}
\label{appendix:spasm}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

The progamming model used by SPASM build up values on the maintained stack
and stores values in specially reserved offsets equivalent to registers 
in order to build up system calls. Immediates can be assigned into VAL
incrementally, and stored, operated on, or pushed onto the maintained
stack. Control Flow is implemented by modifying the maintained 
instruction pointer SRIP, including conditional jmps.

Using this we have implemented various example programs including 
\textit{HelloWorld}, \textit{FizzBuzz}, and \textit{DialbackShell}.
See Figure~\ref{fig:spasm_model} for a diagram of our usage of SPASM 
within the \speculake model.

@@

\item[DADD] (0x0D) - 
Dereference the value in PTR, add VAL and store back into the dereference 
of PTR.

\item[DMUL] (0x0C) - 
Dereference the value in PTR, multiply VAL and store back into the dereference
 of PTR.

\item[DDIV] (0x0B) - 
Dereference the value in PTR, divide by VAL and store back into the 
dereference of PTR.

\item[DSHL] (0x09) - 
Dereference the value in PTR, shift left by VAL and store back into the 
dereference of PTR.

\item[DSHR] (0x08) - 
Dereference the value in PTR, shift right by VAL and store back into the 
dereference of PTR.

\item[DAND] (0x07) - 
Dereference the value in PTR, bitwise and VAL and store back into the 
dereference of PTR.

\item[DOR] (0x06) - 
Dereference the value in PTR, bitwise or with VAL and store back into the 
dereference of PTR.

\item[DXOR] (0x05) - 
Dereference the value in PTR, bitwise xor with VAL and store back into the 
dereference of PTR.

>>>>

\pagebreak
\section{SPASM Instruction Set}
\label{appendix:spasm}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

The progamming model used by SPASM builds up values on the maintained stack
and stores values in specially reserved offsets equivalent to x86 registers 
in order to perform system calls. Immediates can be assigned into VAL
incrementally, and stored, operated on, or pushed onto the maintained
stack. Control Flow, including conditional jmps, is implemented by modifying 
the maintained instruction pointer SRIP.

Using this we have implemented various example programs including 
\textit{HelloWorld}, \textit{FizzBuzz}, and \textit{DialbackShell}.
See Figure~\ref{fig:spasm_model} for a diagram of the SPASM emulator's
usage within the \speculake model.

@@

\item[DADD] (0x0D) - 
Dereference the value in PTR, add VAL and store the result back into the dereference 
of PTR.

\item[DMUL] (0x0C) - 
Dereference the value in PTR, multiply VAL and store the result back into the dereference
 of PTR.

\item[DDIV] (0x0B) - 
Dereference the value in PTR, divide by VAL and store the result back into the 
dereference of PTR.

\item[DSHL] (0x09) - 
Dereference the value in PTR, shift left by VAL and store the result back into the 
dereference of PTR.

\item[DSHR] (0x08) - 
Dereference the value in PTR, shift right by VAL and store the result back into the 
dereference of PTR.

\item[DAND] (0x07) - 
Dereference the value in PTR, bitwise and with VAL and store the result back into the 
dereference of PTR.

\item[DOR] (0x06) - 
Dereference the value in PTR, bitwise or with VAL and store the result back into the 
dereference of PTR.

\item[DXOR] (0x05) - 
Dereference the value in PTR, bitwise xor with VAL and store the result back into the 
dereference of PTR.

### 2018-05-07 paper/architecture.tex
Read through - minor updates

%Speculative branching is not a new feature in processors~\cite{branching-hist}, 
%and it has been integrated in multiple generations of commodity processors today. 
%Accordingly speculative instruction capacity has fairly well defined limits, in 
%line with known features of processor designs. We have determined that features 
%like reorder buffers (ROB), physical register file (PRF), and cache miss timing
%determines a strong limit on the number of instructions that can be accomplished
%speculatively.

\subsubsection{Cache Miss Duration}
When executing instructions speculatively we rely on a memory load of a function
pointer from uncached memory. Thus, one potential limit on our computation comes
in the form of the time it takes for the memory read to return with a result
(and for the CPU to determine the result was mis-predicted).
We measured the number of cycles a
cache miss takes to return by artificially evicting an item from cache and
timing reads from its address.
Figure~\ref{fig:cache-miss} shows the CDF of cycles taken. In the typical case,
an evicted item takes approximately 300 clock cycles to load from L3, which
would allow a limit of roughly 200 speculative instructions to be
executed during that time. We note that when an item is not in L3, it takes
considerably longer to load, in theory allowing for thousands of speculative
instructions.

@@

We also tested whether instruction complexity or data dependencies impact
the number of instructions that can be completed. We find that data dependencies
and instruction complexity both have an impact on the number of instructions
that can be executed. Instruction complexity is determined the number of
$\mu$ ops that the instruction uses, which appears to be what is tracked in the ROB.
For instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59
$\mu$ ops, and we can execute up to 3 of them in the speculative world.
Meanwhile, we can execute up to 18 32-bit \texttt{idiv} instructions, which each
take 9 $\mu$ ops~\cite{intel-instruction-tables}.

We found that the processor can identify
some operations that have no effect on output registers and allows them to use the 
entire size provided by the ROB, as determined in section~\ref{sssec:ROB}. This includes 
idomatic no-ops and zero idioms, which have no reliance on register values, 
for example the instruction \textit{xchg \%rax \%rax} and \textit{xor \%rax, \%rax}. 
However, if an instruction is a potential data hazard, it must use an entry in the PRF
until it reaches the correct stage in the execution pipeline. In this case, the number 
of instructions could be limited by the PRF entries available.

@@

We find that not all Intel CPUs we tried support nested speculation. For
example, it appears Haswell chips do not speculate while already speculating,
but nonetheless support non-nested \speculake. Both Sandy Bridge (which
preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
speculation.
% Could cut
We leverage this support in Section~\ref{subsec:decryption} to
increase the security of speculative decryption keys.

>>>>

\subsubsection{Cache Miss Duration}
When executing instructions speculatively we rely on a memory load of a function
pointer from uncached memory. Thus, one potential limit on our computation comes
in the form of the time it takes for the memory read to return with a result
(and for the CPU to determine the result was mis-predicted).
We measured the number of cycles a
cache miss takes to return by artificially evicting an item from cache and
timing reads from its address.
Figure~\ref{fig:cache-miss} shows the CDF of cycles taken. In the typical case,
an evicted item takes approximately 300 clock cycles to load from L3, which
would allow a limit of roughly 200 speculative instructions to be
executed during that time. We note that when an item is not in L3, it takes
considerably longer to load, in theory allowing for thousands of speculative
instructions in a significant fraction of runs.

@@

We also tested whether instruction complexity or data dependencies impact
the number of instructions that can be completed. We find that data dependencies
and instruction complexity both have an impact on the number of instructions
that can be executed. Instruction complexity is determined by the number of
$\mu$ ops that the instruction uses, which appears to be what is tracked in the ROB.
For instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59
$\mu$ ops, and we can execute up to 3 of them in the speculative world.
Meanwhile, we can execute up to 18 32-bit \texttt{idiv} instructions, which each
take 9 $\mu$ ops~\cite{intel-instruction-tables}.

We found that the processor can identify
some operations that have no effect on output registers and allows them to use the 
entire size provided by the ROB, as determined in section~\ref{sssec:ROB}. This includes 
idomatic no-ops and zero idioms, which have no reliance on register values, 
for example the instruction \texttt{xchg \%rax \%rax} and \texttt{xor \%rax, \%rax}. 
However, if an instruction is a potential data hazard, it must use an entry in the PRF
until it reaches the correct stage in the execution pipeline. In this case, the number 
of instructions could be limited by the PRF entries available.

@@

We find that not all Intel CPUs support nested speculation. For
example, it appears Haswell chips do not speculate while already speculating,
but nonetheless support non-nested \speculake. Both Sandy Bridge (which
preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
speculation.
% Could cut
We leverage this support in Section~\ref{subsec:decryption} to
increase the security of speculative decryption keys.

### 2018-05-07 paper/discussion.tex
Read through - minor updates

\subsubsection{Multiple Triggers}
To create further difficulties for a reverse engineer, or to further target
the execution environment in which a sample executes it is possible
to have the payload program test a series of separate jump patterns. The decryption
process can then be done in stages where the result of a given stage determines 
the initial state of the next decryption stage that the payload program 
performs. Only when all triggers have been completed in the correct order will
the true plaintext be available.

While this would increase the difficulty for a reverse engineer who has access
to a sample in isolation, it  does not add any extra layers of complication 
for reverse engineers who can observe the entire decryption process. Any state 
encoded in the real world during the individual stages will be available to current 
debugging tools. This allows reverse engineers start the sample from the initial
state of any stage in subsequent runs. For example, they could patch the 
binary to begin with the state of the final stage such that only the final 
trigger must be run again to retreive the plaintext. While this simplifies the 
job of acquiring the plaintext of a given \speculake sample, it does not give any 
information about the trigger programs or the key schedules used.

>>>>

\subsubsection{Multiple Triggers}
To create further difficulties for a reverse engineer, or to further target
the execution environment it is possible
to have the payload program test a series of separate jump patterns. The decryption
process can then be done in stages where the result of a given stage determines 
the initial state of the next decryption stage that the payload program 
performs. Only when all triggers have been completed in the correct order will
the true plaintext be available.

While this would increase the difficulty for analyst  who has access
to a sample in isolation, it  does not add any extra layers of complication 
for an analyst who can observe the entire decryption process. Any state 
encoded in the real world during the individual stages will be available to current 
debugging tools. This allows the initial state of a sample to be modified
to simplify subsequent runs. For example, they could patch the 
binary to begin with the state of the final stage such that only the final 
trigger must be run again to retreive the plaintext. While this simplifies the 
job of acquiring the plaintext of a given \speculake sample, it does not give any 
information about the trigger programs or the key schedules used.

### 2018-05-07 paper/introduction.tex
add high-level figure
Related work and reference filling.
Read through - minor updates

Existing malware use several techniques to evade detection and
make it difficult for analysts to determine payload behavior of reported malware. 
For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
data that must be ``unpacked'' at runtime, making it difficult to tell
statically what a program will do~\cite{malware-packers}. Malware may also use
\emph{triggers} that only run the payload when certain conditions are present, preventing
it from executing when it is inside an analysis sandbox or debugger.

@@

While it's possible for the trigger and payload programs to be bundled in a
single program, an analyst aware of \speculake could easily find the target
function in the payload program based on the jump pattern in the trigger program.
However, if the trigger and payload are kept separate, the analyst has a much
harder job and must identify both.
Moreover, it is possible that the trigger program be a
seemingly-unrelated benign program already on the victim's computer. We
investigate using the OpenSSL library as a benign trigger
program in Section~\ref{sec:tk}. If the trigger program is another benign
program on the system, the analyst has the difficult task of identifying which
program, library, or even operating system component is responsible for training
the CPU's branch predictor, and finding the specific set of jumps that occur at
runtime. To make matters worse, the payload program can include dummy jump
patterns that are effectively dead ends for the analyst, as they do not
ultimately call the target function speculatively.

@@

Using these building blocks, we demonstrate the practicality of hiding arbitrary
computation by implementing a reverse shell in our speculative instruction set.
We show that this simple payload is able to perform several system calls
in a reasonable time, ultimately launching a dial-back TCP shell in under 8~seconds after
the trigger program starts.

>>>>

Existing malware use several techniques to evade detection and
make it difficult for analysts to determine payload behavior of reported malware. 
For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
data that must be ``unpacked'' at runtime, making it difficult to tell
statically what a program will do~\cite{malware-packers}. Malware may also use
\emph{triggers} that only run the payload when certain conditions are present, preventing
it from executing when it is inside an analysis sandbox or 
debugger~\cite{balzarotti2010efficient,red-pill}.

@@

%Figure~\ref{fig:high-level}

\FigHighLevel

While it's possible for the trigger and payload programs to be bundled in a
single program, an analyst aware of \speculake could easily find the target
function in the payload program based on the jump pattern in the trigger program.
However, if the trigger and payload are kept separate, the analyst has a much
harder job and must identify both.
Moreover, it is possible that the trigger program be a
seemingly-unrelated benign program already on the victim's computer. We
investigate using the OpenSSL library as a benign trigger
program in Section~\ref{subsec:openssl}. If the trigger program is another benign
program on the system, the analyst has the difficult task of identifying which
program, library, or even operating system component is responsible for training
the CPU's branch predictor, and finding the specific set of jumps that occur at
runtime. To make matters worse, the payload program can include dummy jump
patterns that are effectively dead ends for the analyst, as they do not
ultimately call the target function speculatively.

@@

Using these building blocks, we demonstrate the practicality of hiding arbitrary
computation by implementing a reverse shell in our speculative instruction set.
We show that this simple payload is able to perform several system calls
in a reasonable time, ultimately launching a dial-back TCP shell in under 7~seconds after
the trigger program starts.

### 2018-05-07 paper/payloads.tex
Read through - minor updates

\subsection{Turing Machine}
\label{subsec:turing}
To demonstrate that arbitrary computation can be done in the speculative world
we implement a Turing machine, configured to run the 5-state Busy Beaver 
function~\cite{chaitin1987computing,herken1992universal}.
The Busy Beaver is a configuration that writes the most
1s on the tape with only a limited number of states and still halts.

@@

Section~\ref{subsec:turing} demonstrates that arbitrary speculative computation
is possible, but hiding malware this way has several drawbacks. First, an
analyst might be able to observe that something is driving a Turing
machine by noticing the real world state updating. With enough symbols, the
analyst could reverse engineer the state machine even without being able to see
the internal logic executing speculatively. Second, Turing machines are a
poor choice for practical computation, as they are inefficient and have no direct
way to perform system calls or otherwise interface externally.

We explore a more practical application of using \speculake to perform
decryption speculatively. To hide keys from the analyst, the key and decryption
code only occur in the speculative world, while committed state only learns the
resulting plaintext. Furthermore, the speculative code can
derive the decryption key based entirely off the trigger program, making it
impossible for an analyst that only has the payload program from being able to
determine the key or decrypt the embedded ciphertext.

@@

\subsection{Virtual Machines}
We demonstrated in Section~\ref{subsec:turing} that general purpose computation 
is possible in the speculative environment in spite of the contraints 
imposed by the speculative world. However, a turing machine
is not an efficient or usable model for computation when
attempting to accomplish tasks on modern processors.

@@

To accommodate both developers as well as the unique constraints 
enforced by the speculative primitive, a custom emulation solution is required. We 
have approached this challenge from both directions. First we demonstrate 
general purpose emulation using a custom instruction set -- SPASM -- which
accommodates the speculative primitive specifically. Next we investigate custom 
wrappers on traditional emulators allow for similar payload construction,
with a much more developer friendly

>>>>

\subsection{Turing Machine}
\label{subsec:turing}
To demonstrate that arbitrary computation can be done in the speculative world
we implement a Turing machine, configured to run the 5-state Busy Beaver 
function~\cite{chaitin1987computing,herken1992universal}.
The Busy Beaver is a configuration which writes the most
1s on the tape before halting with a limited number of states.

@@

While a Turing machine demonstrates that arbitrary speculative computation
is possible, hiding malware this way has several drawbacks. First, an
analyst might be able to observe that something is driving a Turing
machine by noticing the real world state updating. With enough symbols, the
analyst could reverse engineer the state machine even without being able to see
the internal logic executing speculatively. Second, Turing machines are a
poor choice for practical computation, as they are inefficient and have no direct
way to perform system calls or otherwise interface externally.

We explore a more practical application of using \speculake to perform
decryption speculatively. To hide keys from the analyst, the key and decryption
code only occur in the speculative world, while committed state only learns the
resulting plaintext. Furthermore, the speculative code can
derive the decryption key based entirely off the trigger program, making it
impossible for an analyst that only has the payload program to
determine the key or decrypt the embedded ciphertext.

@@

\subsection{Virtual Machines}
We demonstrate in Section~\ref{subsec:turing} that general purpose computation 
is possible in the speculative environment in spite of the contraints 
imposed by the speculative world. However, a turing machine
is not an efficient or usable model for computation when
attempting to accomplish tasks on modern processors.

@@

To accommodate both developers as well as the unique constraints 
enforced by the speculative primitive, a custom emulation solution is required. 
we have approached this challenge by creating a custom instruction set -- 
SPASM -- which accommodates the speculative primitive specifically. This 
provides a higher level of abstraction around the speculative primitive
that we have used to implement various payloads.

% The assembly level abstraction that SPASM provides is still not extremely
% user friendly as it does not allow an author to write their encrypted 
% emulator payload in a high level language.

### 2018-05-07 paper/related-work.tex
Read through - minor updates

Many resourcess have been useful in developing the \speculake exploit 
model.

\subsection{Weird Machines \& Red-Pills}
Hardware side effects often leave behind unexpected state transitions 
that are not, or cannot be, accounted for in modeling environments. Malware
has historically made use of these bugs to create "weird" models of 
computation~\cite{weird_machines}. These state transitions can also leak 
information about the environment in which a program is running, which 
malware has used to detect honey-pots and emulated environments.

@@

Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
can be constructed using the page fault handling mechanism in Intel's
IA32 architecture~\cite{bangert2013page}. Bugs like this are abundant in
modern processors as multiple threads share system resources cooperatively
and optimizations is done objectively with regards of the 
process isolation for the most part. Accessing copoerative hardware resources is at
the heart of the \speculake model, which constructs a hardware weird machine
using speculative branching and \texttt{Flush+Reload}. This is 
very much in the spirit of traditional exploit creation~\cite{weird_exploits}.

@@

Unlike other red-pills based on hardware bugs, \speculake cannot be bypassed
by patching binaries as the speculation can only happen on "bare-metal" systems, 
and speculative execution will not trigger debugging break-points as 
they are never completed in a committed state.
Symbolic execution will also fail to find a target function as it is never on the 
execution pathway.

@@

The \speculake model relies critically on cache side channel attacks to 
communicate information from the speculative world. Cache side-channels have 
been used against a number of processors at essentially all different 
caching levels -- L1 data cache~\cite{percival2005cache,zhang2012cross,osvik2006cache}, 
L1 instruction cache~\cite{aciiccmez2010new}, 
L2 cache~\cite{ristenpart2009hey,percival2005cache}, 
LLC~\cite{ristenpart2009hey,liu2015last}, 
and branch prediction cache~\cite{aciiccmez2007power}. 
These attacks have also been performed in 
cloud~\cite{ristenpart2009hey,zhang2012cross}, 
browser~\cite{oren2015spy,google_cache_browser},
security critical encryption~\cite{yarom2014recovering,tromer2010efficient}
environments, demonstrating how ubiquitous hardware side-channel exploits can be.

@@

% Lead in 
Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
multiple teams were working to reverse engineer the mechanisms that implemented in 
the microcode of various processors and document their behavior~\cite{intel-instruction-tables, uop_article}.
Efforts to examine cache and memory security also preempted the spectre work, 
including Sophia D'Antoine's thesis work developing a side channel based on 
instruction interleavings in the CPU~\cite{d2015exploiting}. Similar research into branch 
predictors, cache replacement policies, and reorder buffer construction including work by Matt 
Godbolt~\cite{godbolt2016branch}, Henry Wong~\cite{measuring-rob} paved the runway for 
dedicated branch prodiction side-channels.

% Discovery & Followup
Given the numerous directions that were being investigated in reverse engineering 
Intel processors, multiple teams discovered the Spectre vulnerabilites concurrently including the 
various teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
at Google~\cite{project_zero},and various others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
These works propose multiple variants making use of banch prediction.  Follow 
up work on speculative execution side channels have also demonstrated
web-based vulnerabilites~\cite{genkin2018drive} as well as hardware isolation  
leaks in Intel's SGX~\cite{spectre_sgx}.

% defenses 
Much like cache side channel vulnerabilities speculative execution is relatively
critical to processor performance and requires significant software, and hardware 
updates to fully address. Multiple measures have been proposed to address the 
individial variants of Spectre.  Trippel \textit{et al.} proposes a method of faithfully simulating 
hardware bugs, such as processes cooperatively using speculating branch prediction,
by synthesizing microarchetecture defects~\cite{trippel2018meltdownprime}. However, 
manufacturers generally guard complete formal descriptions of processor microarchitecture closely, 
making this a less accessible solution for the research community.

@@

>>>>

% Many resourcess have been useful in developing the \speculake exploit 
% model.

\subsection{Weird Machines \& Red-Pills}
Hardware side effects often leave behind unexpected state transitions 
that are not, or cannot be, accounted for in modeling environments. Malware
has historically made use of these bugs to create "weird" models of 
computation~\cite{weird_machines}. These state transitions can also leak 
information about the environment in which a program is running, which 
malware has used to detect honey-pots and emulated 
environments~\cite{balzarotti2010efficient}.

@@

Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
can be constructed using the page fault handling mechanism in Intel's
IA32 architecture~\cite{bangert2013page}. Bugs like this are abundant in
modern processors~\cite{d2015exploiting} as multiple threads share system resources cooperatively
and optimizations is done objectively with regards of the 
process isolation for the most part~\footnote{With the exception of permission 
rings, i.e. kernel vs user space isolation}. Accessing cooperative hardware resources is at
the heart of the \speculake model, which constructs a hardware weird machine
using speculative branching and \texttt{Flush+Reload}. This is 
very much in the spirit of traditional exploit creation~\cite{weird_exploits}.

@@

Unlike other red-pills based on hardware bugs, \speculake cannot be bypassed
by patching binaries as the speculation can only happen on "bare-metal" systems, 
and the critical functionality is performed speculatively. Similarly, debugging 
break-points wll not be triggered as instructions are never completed in a committed state.
Symbolic execution will also fail to find a target function as it is never on the 
execution pathway.

@@

The \speculake model relies critically on cache side channel attacks to 
communicate information from the speculative world. Cache side-channels have 
been used against a number of processors at essentially all different 
caching levels -- L1 data cache~\cite{percival2005cache,zhang2012cross,osvik2006cache}, 
L1 instruction cache~\cite{aciiccmez2010new}, 
L2 cache~\cite{ristenpart2009hey,percival2005cache}, 
LLC~\cite{ristenpart2009hey,liu2015last}, 
and branch prediction cache~\cite{aciiccmez2007power}. 
These attacks have also been performed in 
cloud~\cite{ristenpart2009hey,zhang2012cross}, 
browser~\cite{oren2015spy,google_cache_browser},
and security critical encryption~\cite{yarom2014recovering,tromer2010efficient}
environments, demonstrating how ubiquitous hardware side-channel exploits can be.

@@

% Lead in 
Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
multiple teams were working to reverse engineer the mechanisms that implemented in 
the microcode of various processors and document their behavior~\cite{intel-instruction-tables, uop_article}.
Efforts to examine cache and memory security also preempted the spectre work, 
including Sophia D'Antoine's thesis work developing a side channel based on 
instruction interleavings in the CPU~\cite{d2015exploiting}. Similar research into branch 
predictors, cache replacement policies, and reorder buffer construction including work by Matt 
Godbolt~\cite{godbolt2016branch}, and Henry Wong~\cite{measuring-rob} paved the runway for 
dedicated branch prodiction side-channels.

% Discovery & Followup
Given the numerous directions that were being investigated in reverse engineering 
Intel processors, multiple teams discovered the Spectre vulnerabilites concurrently including 
the teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
at Google~\cite{project_zero},and various others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
These works propose multiple variants making use of banch prediction.  Follow 
up work on speculative execution side channels have also demonstrated
web-based vulnerabilites~\cite{genkin2018drive} as well as hardware isolation  
leaks in Intel's SGX~\cite{spectre_sgx}.

% defenses 
Much like cache side channel vulnerabilities speculative execution is relatively
critical to processor performance and requires significant software, and hardware 
updates to fully address. Multiple measures have been proposed to combat the 
individial variants of Spectre.

@@

Trippel \textit{et al.} proposes a method of faithfully simulating 
hardware bugs, such as processes cooperatively using speculating branch prediction,
by synthesizing microarchetecture defects~\cite{trippel2018meltdownprime}. However, 
manufacturers generally guard complete formal descriptions of processor microarchitecture closely, 
making this a less accessible solution for the research community.
